\sysfull

### 2018-04-12 paper/systemdescription.tex
re-wrote the overview part

\sys is a new programming and execution model for intermittent computing on energy-harvesting devices. \sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em accessible}, {\em efficient} and {\em flexible}. \sys accomplishes this goal with a constellation of a new programming model, compiler, and run time software system support, refer to Figure~\ref{fig:system_overview}.

@@

\textbf{\sys Programming Model.} To use \sys, a programmer first writes plain, imperative C code. The programmer then has an option to {\em manually} or {\em automatically} translate the program into code for \sys's task-based programming model. To manually translate, the programmer decomposes the program into tasks, and annotates memory accesses that manipulate data shared by multiple tasks. To automatically translate, the programmer simply uses \sys's compiler (described in Section~\ref{sec:compiler}), which decomposes a program into tasks and annotates accesses to task-shared data. After translating to task-based code and compiling, the programmer has an executable \sys binary.

\sys's programming interface, listed in Table~\ref{tab:viper_syntax}, consists of
constructs for defining tasks~\cite{chain,alpaca} and accessors for
protected variables shared between tasks. These anotations are added either by
the programmer (in the manual translation mode) or the compiler (in automatic
translation mode). As in prior task-based intermittent programming
models~\cite{chain,alpaca}, tasks are functions that are not allowed to have
any direct callers but receive control by explicit {\tt next\_task} statements.
In \sys, a task transition atomically commits the memory operations of the
current task.  The programmer identifies each task by adding the {\tt task}
keyword to the function's declaration and designates one task as the {\em
origin} task at which execution begins. After a power failure,
\sys restarts from the last (partially) executed task.
%
A \sys program manipulates {\em protected} state---accessed by more than one
task---using the \emph{read variable} ({\tt RVAR}) and \emph{write variable}
({\tt WVAR}) operations. These operations convey to \sys's memory manager which
data must be buffered in volatile memory and committed.

@@

\textbf{\sys Memory Manager.} At runtime \sys's \emph{virtualizing memory
manager} (Section~\ref{sec:memory_virtulaization}) protects changes to
non-volatile data against becoming inconsistent after a power failure.  Relying
on annotations of accesses to protected variables, \sys's memory manager {\em
pages} these data into volatile memory from non-volatile memory as they are
needed.  Tasks cannot modify protected variables directly in non-volatile
memory.  If the working set exceeds the size of volatile memory, pages are
swapped out to non-volatile memory without committing the changes, leaving
original locations of protected variables unchanged.  When a coalesced task
completes, \sys atomically commits modified pages to their original locations
in non-volatile memory.

\textbf{\sys Task Scheduler.} The scheduler {\em coalesces}
statically\hyp{}defined tasks into dynamic tasks sized to match the available
energy. By default, tasks run in sequence and each task commits as it
completes, potentially incurring unnecessary overhead for commits between
consecutive, non-failing tasks. \sys dynamically coalesces consecutive tasks by
deferring the commit of an earlier task until the completion of a later task.
\sys's memory management mechanism is essential for coalescing, because commits
are deferred by keeping dirty pages in volatile memory across coalesced
boundaries.  After a coalesced sequence of tasks completes, \sys commits pages
updated by all of the tasks to non-volatile memory, reducing the commit
overhead through batching.

>>>>

\sys is a new programming and execution model for intermittent computing on
energy-harvesting devices. \sys addresses the challenges outlined in
Section~\ref{sec:background} to make task-based intermittent programs {\em
programmable} and {\em efficient}. \sys accomplishes this goal with a
constellation of a new programming model, compiler, and run time software
system support, that supports dynamically adaptive task-based execution.
Figure~\ref{fig:system_overview} shows an overview of \sys.

@@

\textbf{\sys Programming and Execution Model.} 
To use \sys, a programmer first writes plain, imperative C code. The programmer
then must decompose the program into tasks. To manually decompose a program
into tasks, the programmer designates a set of functions as tasks, sequences
control-flow between these tasks, and annotates memory accesses that manipulate
data shared by multiple tasks.

\sys's programming interface, listed in Table~\ref{tab:viper_syntax}, consists
of primitives for defining tasks~\cite{chain,alpaca} and accessors for
protected variables shared between tasks. As in prior task-based intermittent
programming models~\cite{chain,alpaca}, tasks are functions that are not
allowed to have any direct callers but receive control via explicit {\tt
next\_task} statements only.  The programmer identifies each task
by adding the {\tt task} keyword to the function's declaration and designates
one task as the {\em origin} task at which execution begins. After a power
failure, \sys restarts the last partially executed task.
%
A \sys program manipulates {\em protected} state---accessed by more than one
task---using the \emph{read variable} ({\tt RVAR}) and \emph{write variable}
({\tt WVAR}) operations. These operations convey to \sys that
the referenced data are task-shared, meaning that they must be buffered in
volatile memory and committed when the executing task ends.
%
Commit is a key overhead associated with task-based operation and \sys's
coalescing strategy (described in Section~\ref{sec:coalescing} avoids commit
when a task ends by continuing to buffer task-shared variables for multiple
coalesced tasks, and committing only when all tasks complete.

@@

The programmer can decompose a program into tasks either {\em manually} or {\em
automatically}. To do so manually,  requires reasoning similar to prior
task-based systems~\cite{chain,alpaca}.  
%
A programmer may also opt to use compiler support to automatically decompose a
program into tasks, leveraging recent work~\cite{cleancut,baghsorkhi_cgo_2018}.
%
Without loss of generality, we assume throughout this work that the programmer
manually decomposed the program into tasks; \sys's behavior with automatically
decomposed code would be identical.
%

Regardless of whether automatically or manually decomposed into tasks, the
programmer compiles their task-based code, and links to the \sys runtime,
producing a \sys-enabled binary.
%
The \sys runtime implements \sys's task-based programming and execution model.
%
The runtime system also includes \sys's novel {\em virtualizing memory manager}
and its {\em task coalescing manager}, both of which are essential to \sys's
adaptive task-based execution.

\textbf{\sys Task Coalescing Manager.} \sys relies on its task coalescing
manager to dynamically {\em coalesce} statically\hyp{}defined tasks to avoid
inessential overheads associated with completing tasks.  
%
By default, tasks run in a sequence and each task commits its task-shared state
as it completes.  The time and energy cost of a task's commit is unnecessary if
the task and its successor both complete without a power failure.  The key
insight that \sys leverages is that the first task could have deferred its
commit to the second task. {\em coalescing} tasks by deferring the first commit
avoids the fixed cost of the first commit (task transitioning overheads).
Coalescing also amortizes per-variable commit cost, committing each  location
accessed by both tasks only after the second task, rather than once per task.
%

\textbf{\sys Virtual Memory Manager.} 
\sys is able to efficiently coalesce tasks because of its efficient virtual
memory manager,
%
which is described in Section~\ref{sec:memory_virtualization}. 
%
\sys's memory manager paginates memory and ensures that data in a page remain consistent
despite power interruptions. 
%
\sys allows a task to manipulate data in a volatile copy of a page only.
%
Pages swap between volatile and non-volatile memory, depending on the capacity
of the volatile memory and the program's access pattern. 
%
\sys tracks a task's memory accesses efficiently at page granularity (rather
than using, e.g., costly word-granular tracking).
%
When a task ends, each page it accessed commits from volatile memory (or from a
non-volatile swap region for dirty pages) back to the non-volatile main memory.
%
Pages efficiently, atomically commit using a two-phase commit procedure
accelerated using hardware support for direct memory access (DMA). 
%
